Tampering Detection In Low-Power Smart Cameras

### 2015-05-28 paper.tex
minor edits

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampering attempt. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, blurring detection, displacement detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as in Fig. \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig. \ref{fig:neve}. Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes. We refer to these events/attacks as tampering. In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.

@@

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts. In fact, tampering prevents the correct scene interpretation: even a mild blur might hinder the identification of important details such as licence plates. Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g. around few frames per second): in these conditions, several tampering-detection algorithms have been presented.

In particular, camera displacement and occlusions are typically detected by background models or frame comparison, while blurring is instead detected by monitoring the high-frequency component of images, as this is expected to drop. In ~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain. Two background models computed over different time intervals are used in~\cite{saglam2009real}. In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation. Background models and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while number of SURF~\cite{bay2006surf} keypoints is used to detect blurring. A frame buffer rather than an explicit background is used~\cite{ribnick2006real}. All these solutions, including~\cite{ribnick2006real,harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.

@@

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, while being more challenging than in conventional surveillance cameras~\cite{perrig2004security}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations \cite{piccardi2004background}. In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}. Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames. Moreover, false alarms represent a serious problem in low-power smart cameras as these result in useless data processing and network transmissions, thus energy waste.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}. The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity is used to detect displacement, and the average gradient norm to detect blurring. These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera. In particular, we show that separately monitoring the indicators over different regions of the image can substantially improve the displacement-detection performance. Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image. Thus, our solution represents as a prompt trigger that can be possibly combined with other sequential monitoring techniques. Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect time $\tau^*$ when any tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs. We assume that $T_0$ tampering-free frames and provided for training.

@@

Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$). The former is meant to detect camera displacement, the latter camera defocus (because blur cases a drop in the high-frequency content). As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section \ref{subsec:Segmentation}.

@@

\subsubsection{Computational Complexity}

@@

%
%% \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
%
%
%\subsection{Performance Assessment}
%\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
%
%\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
\begin{figure}[t]
\centering
\subfigure[]{\label{fig:ROCdisplacement}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement}}
\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus}}
\caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
\label{fig:ROC}
\end{figure}
%\begin{figure}[htb]
%\centering
%\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
%\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
%\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
%\label{fig:ROCdisplacement}
%\label{fig:ROCdefocus}
%\end{figure}
Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blur/defocus, it is more convenient to consider the whole scene at once.

\begin{table}[tbh]
\centering
\begin{tabular}{l|l|l|l|l|l|l|}
\cline{2-7}
& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Defocus Detection}} \\ \cline{2-7} 
& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
\end{tabular}
\caption{Area under curve of the analized solutions}
\label{tab:AUC}
\end{table}

>>>>

We here introduce a tampering-detection algorithm specifically designed for low-power smart cameras: the algorithm leverages very simple indicators that are then monitored by an outlier-detection scheme. Any frame yielding anomalous indicator is detected as a tampered frame. Core of the algorithm is the partitioning of the scene into adaptively defined regions, that are preliminarily defined by segmenting the image during the algorithm-configuration phase, and which shows to substantially improve the detection of camera displacements. Our experiments show that the proposed algorithm can successfully operate on sequences acquired at very low-frame rate, such as one frame every minute, with a very small computational complexity. %and leveraging the image partitioning into regions yields improved performance with respect to monitoring the whole scene
\keywords{tampering detection, blurring detection, displacement detection}
\end{abstract}

\section{Introduction}\label{sec:introduction}
% What is the problem?
When cameras operate outdoor and in harsh environments, dust, rain drops or snow flakes might lie on the camera lens resulting in blurry pictures, as shown in Figure \ref{fig:pioggia}, or in partial occlusions of the scene, as in Fig \ref{fig:neve}. Similarly, intentional attacks like displacing the camera, changing its focus, or spraying some opaque or glossy liquid over the lenses, would result in heavily compromised pictures, that would be useless for monitoring purposes. We refer to these events/attacks as tampering. In some cases, tampering is easy to detect, e.g., when the camera integrity is affected and the device goes out-of-order. However, in many other situations, namely when the device is not physically damaged, tampering detection is not straightforward, and sometimes image analysis is the only viable option.

@@

% Why is it interesting and important?
Tampering detection is an essential feature in surveillance systems~\cite{hampapur2005smart}, where cameras are expected to autonomously detect any tampering, and promptly report alerts. In fact, tampering prevents the correct interpretation of the scenex: even a mild blur might hinder the identification of important details such as licence plates. Surveillance cameras are typically connected to the power supply and operates at normal frame-rates (e.g. around few frames per second): in these conditions, several tampering-detection algorithms have been presented.

In particular, camera displacement and occlusions are typically detected by estimating the background or computing frame difference, while blurring is instead detected by monitoring the high-frequency components of images, as these are expected to drop. In~\cite{aksay2007camera} and~\cite{saglam2009real} tampering is detected by comparing histograms of background and current frame, and by monitoring the energy in wavelet or Fourier domain. Two background models, estimated over different time intervals, are used in~\cite{saglam2009real}. In~\cite{gil2007automatic}, tampering detection is performed by combining background subtraction together with edge detection and normalized cross-correlation. Background estimates and block matching in~\cite{tsesmelis2013tamper} enable the displacement detection, while the number of SURF~\cite{bay2006surf} keypoints is used to detect blurring. A frame buffer rather than an explicit background is used~\cite{ribnick2006real}. All these solutions, including~\cite{harasse2004automated,kryjak2012fpga} are computationally demanding, and are not viable options for low-power cameras.

@@

% Why is it hard? (E.g., why do naive approaches fail?)
Tampering detection in low-power smart cameras has not been much investigated in the literature~\cite{alippi2010detecting}, also because it is more challenging than in conventional surveillance cameras~\cite{perrig2004security}. Beside computational aspects -- such as the number of operations per pixels allowed -- the big issue is that low-power smart cameras typically operate at very low-frame rates (e.g., less than one frame per minute), thus the acquired sequence does not evolve smoothly. This prevents the use of learned background models and the analysis of foreground variations. In fact, when dynamic environments are acquired at low frame rates, two consecutive frames might be very different because of changes in the scene and in the light conditions, yielding sequences like those in Figure~\ref{fig:sequences}. Smart cameras have to promptly distinguish between \emph{normal} scene or illumination changes, and changes due to camera tampering to avoid the transmission of corrupted frames. Moreover, false alarms represent a serious problem in low-power smart cameras since they result in energy waste because of useless data processing and network transmissions.
% We here address the problem of detecting tampering events on a camera device that is employed for monitoring purpose. Tampering events can be due to malicious attacks, as for instance a displacement of the camera or spraying some dirty liquid that would prevent the acquisition of (part of) the monitored scene or the interpretation of the content (e.g., the identification of licence plates).

%What are the key components of my approach and results? Also include any specific limitations.
We address the detection of camera blurring and displacement, which are presented in Section \ref{sec:probForm}. The proposed algorithm, detailed in Section~\ref{sec:propSol}, relies on two indicators: the average image intensity is used to detect displacement, and the average gradient norm to detect blurring. These indicators can be easily computed and monitored by an outlier-detection technique in a low-power smart camera. In particular, we show that separately monitoring the indicators over different regions of the image can substantially improve the displacement-detection performance. Remarkably, these regions can be preliminarily defined during an initial configuration phase (Section \ref{subsec:Segmentation}), thus the algorithm operates at a negligible computational overhead w.r.t. monitoring the whole image (Section \ref{subsec:coputationalComplexity}). Thus, our solution represents a prompt trigger that can be possibly combined with other sequential monitoring techniques. Experiments in Section \ref{sec:experiments} show that leveraging image regions can substantially improve the displacement-detection performance. % Concluding remarks and discussions are given in Section \ref{sec:Conclusion}.

@@

The proposed tampering-detection algorithm analyzes a sequence of frames $\{z_t\}_t$ to detect the time instant $\tau^*$ when tampering like \eqref{eq:model_defocus} or \eqref{eq:model_displacement} occurs. We assume that $T_0$ tampering-free frames and provided for training.

@@

Algorithm \ref{alg:DISPL} presents the proposed tampering-detection algorithm, which relies on two indicators that are very simple to compute: the average intensity (also referred to as \emph{luma}, and denoted by $l$) and the average norm of the gradient (denoted by $g$). The former is meant to detect camera displacement, the latter blurring over camera lens (because blur cases a drop in the high-frequency content). As anticipated in Section~\ref{sec:introduction}, during the initial configuration, the scene is segmented in $K$ disjoint regions $\{R_k, k = 1,\dots,K\}$, namely $R_k \subset \mathcal{X}, R_i \cap R_j = \emptyset, \forall i \neq j$. The employed segmentation is detailed in Section \ref{subsec:Segmentation}.

@@

\subsection{Computational Complexity}\label{subsec:coputationalComplexity}

@@

%
%% \subsection{Alternative Approaches}\label{subsec:AlternativeApproaches}
%
%
%\subsection{Performance Assessment}
%\gi{Adriano: Dire come vengono calcolate le ROC curves TPR e FPR, le cifre di merito insomma, spiegando bene che parametro varia}
%
%\gi{Adriano: metti entrambe le ROC curves, affiancate e per bene ed alcuni esempi di sequenze}
\begin{figure}[t]
\centering
\subfigure[]{\label{fig:ROCdisplacement}\includegraphics[width=0.45\linewidth]{Immagini/ROCdisplacement}}
\subfigure[]{\label{fig:ROCdefocus} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus}}
\caption{ROC curves for displacement detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean luma energy. \textbf{(b)} Analysis of the frame differencing.}
\label{fig:ROC}
\end{figure}
%\begin{figure}[htb]
%\centering
%\subfigure[]{\label{fig:buenosAiresDef}\includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus1}}
%\subfigure[]{\label{fig:buenosAiresDispl} \includegraphics[width=0.45\linewidth]{Immagini/ROCdefocus_fd}}
%\caption{ROC curves for defocus detection, considering three alternative approaches. \textbf{(a)} Analysis of the mean gradient energy. \textbf{(b)} Analysis of the frame differencing.}
%\label{fig:ROCdisplacement}
%\label{fig:ROCdefocus}
%\end{figure}
Experimental results are shown in Figures \ref{fig:ROCdisplacement} and \ref{fig:ROCdefocus}.
As we could see there is an improvement, in the detection of camera displacements, with our approach which separately analizes the behavior of indicators in adaptive regions.
On the other side, as illustrated in Figure \ref{fig:ROCdefocus}, when monitoring indicators meant to detect blurring, it is more convenient to consider the whole scene at once.

\begin{table}[tbh]
\centering
\begin{tabular}{l|l|l|l|l|l|l|}
\cline{2-7}
& \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Displacement Detection}}  & \multicolumn{3}{l|}{\cellcolor[HTML]{C0C0C0}\textbf{Blurring Detection}} \\ \cline{2-7} 
& \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} & \cellcolor[HTML]{EFEFEF}\textbf{Regions} & \cellcolor[HTML]{EFEFEF}\textbf{Total} & \cellcolor[HTML]{EFEFEF}\textbf{Voronoi} \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Luma}}     & 0.9994 & 0.9989 & 0.9985  &            &            &             \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{Gradient}} &  		 &  		  &             & 0.9895 & 0.9996 & 0.9996  \\ \hline
\multicolumn{1}{|l|}{\cellcolor[HTML]{EFEFEF}\textbf{FD}}         & 0.9974 & 0.9944 & 0.9974  & 0.5526 & 0.5322 & 0.5391  \\ \hline
\end{tabular}
\caption{Area under curve of the analized solutions}
\label{tab:AUC}
\end{table}
